---
title: "A logistic + OLS model of water/octanol partition coefficients"
output: html_notebook
---

First we will load in the data for the full set of molecules. The "Nr", "Smiles", "CAS" and "dGow" column is taken from the SI of the paper ["Automated Parametrization of the Coarse-Grained Martini Force Field for Small Organic" Molecules](http://pubs.acs.org/doi/suppl/10.1021/acs.jctc.5b00056) and the other columns have been filled in with the `collect_props.py` Python script.

```{r}
library(readxl)
data <- read_excel("../data.xlsx")
colnames(data)
```

The "Nr", "Smiles", "CAS" and "dGow" was taken from the SI of a published paper, and the other columns were filled in with a Python script.

Next, we will exclude the molecules that are also in the Minnesota valdiation test set.

```{r}
ii<-read.table("../overlap.dat")
data<-data[-ii$V1,]
```

Let's try to do a logistic regression on being positive!. I.e. we will only try to predict the sign.

```{r}
data$sign<-as.integer(data$dGow>0)
gg<-glm(sign~Weight+Heavies+Carbons+Polar+Rings+HDonors+HAcceptors,'binomial',data=data)
summary(gg)
selg<-step(gg,trace=F)
print(summary(selg))
par(mfrow=c(2,2))
plot(selg)
```

Plot a dose response

```{r}
logit<-function(z) {ll<-log(z/(1-z))}
plot(logit(selg$fit),data$sign,xlab="linear predictor",ylab="sign")
lines(sort(logit(selg$fit)),selg$fit[sort.list(selg$fit)])
```

And check its classification

```{r}
pp<-predict(selg,type="response")
pp[pp<.5]<-0
pp[pp>=.5]<-1
table(pp,data$sign)
print(c("Misclassification error=", round(sum(pp!=data$sign)/length(data$sign),3)))
```

Plot the data anew but separate positive and negative values

```{r}
data_pos<-data[data$dGow>0,]
data_pos$Weight<-log10(data_pos$Weight)
data_pos[data_pos$Heavies>15,]$Heavies<-15
data_pos[data_pos$Polar>8,]$Polar<-8
p1<-ggplot(data=data_pos, aes(x=Weight, y=dGow)) + geom_point() 
p2<-ggplot(data=data_pos, aes(x=Heavies, y=dGow)) + geom_point()
p3<-ggplot(data=data_pos, aes(x=Carbons, y=dGow)) + geom_point() 
p4<-ggplot(data=data_pos, aes(x=Polar, y=dGow)) + geom_point() 
p5<-ggplot(data=data_pos, aes(x=Rings, y=dGow)) + geom_point()
p6<-ggplot(data=data_pos, aes(x=HDonors, y=dGow)) + geom_point() 
p7<-ggplot(data=data_pos, aes(x=HAcceptors, y=dGow)) + geom_point() 
library(grid)
library(gridExtra)
grid.arrange(p1, p2, p3, p4, p5, p6, p7, ncol=3)
```

and negative data

```{r}
data_neg<-data[data$dGow<0,]
data_neg$Weight<-log10(data_neg$Weight)
p1<-ggplot(data=data_neg, aes(x=Weight, y=dGow)) + geom_point() 
p2<-ggplot(data=data_neg, aes(x=Heavies, y=dGow)) + geom_point()
p3<-ggplot(data=data_neg, aes(x=Carbons, y=dGow)) + geom_point() 
p4<-ggplot(data=data_neg, aes(x=Polar, y=dGow)) + geom_point() 
p5<-ggplot(data=data_neg, aes(x=Rings, y=dGow)) + geom_point()
p6<-ggplot(data=data_neg, aes(x=HDonors, y=dGow)) + geom_point() 
p7<-ggplot(data=data_neg, aes(x=HAcceptors, y=dGow)) + geom_point() 
library(grid)
library(gridExtra)
grid.arrange(p1, p2, p3, p4, p5, p6, p7, ncol=3)
```

Create an OLS model for the positive molecules

```{r}
mm<-lm(dGow~Weight+Heavies+Carbons+Polar+Rings+HDonors+HAcceptors,data=data_pos)
summary(mm)
par(mfrow=c(2,2))
plot(mm)
```

And then on the negative onese

```{r}
mm<-lm(dGow~Weight+Heavies+Carbons+Polar+Rings+HDonors+HAcceptors,data=data_neg)
summary(mm)
par(mfrow=c(2,2))
plot(mm)
```

Lets try prediction on Minnesota data

```{r}
sota<-read_excel("../phys/minnesota.xlsx")
sota$sign<-as.integer(sota$dGow>0)
pp2<-predict(selg, sota[c("Weight","Carbons","Polar","HDonors","HAcceptors")], type="response")
pp2[pp2<.5]<-0
pp2[pp2>=.5]<-1
table(pp2,sota$sign)
print(c("Misclassification error=", round(sum(pp2!=sota$sign)/length(sota$sign),3)))
```

