---
title: "An OLS model of water/octanol partition coefficients"
output: html_notebook
---

First we will load in the data for the full set of molecules. The "Nr", "Smiles", "CAS" and "dGow" column is taken from the SI of the paper ["Automated Parametrization of the Coarse-Grained Martini Force Field for Small Organic" Molecules](http://pubs.acs.org/doi/suppl/10.1021/acs.jctc.5b00056) and the other columns have been filled in with the `collect_props.py` Python script.

```{r}
library(readxl)
data <- read_excel("../data.xlsx")
colnames(data)
```

The "Nr", "Smiles", "CAS" and "dGow" was taken from the SI of a published paper, and the other columns were filled in with a Python script.

Next, we will exclude the molecules that are also in the Minnesota valdiation test set.

```{r}
ii<-read.table("../overlap.dat")
data<-data[-ii$V1,]
```

Let's look into the colinearity between the molecular properties

```{r}
library(ggplot2)
library(reshape2)

cormat<-melt(cor(data[,seq(5,dim(data)[2])]))

ii<-which(cormat[,"value"]==1)
maxv<-max(cormat[-ii,"value"])
maxi<-which(cormat["value"]==maxv)[1]

ggplot(data = cormat, aes(x=Var1, y=Var2, fill=value)) + scale_fill_gradient2(low="green", high="red", mid="white", midpoint=0, limit=c(-1,1), space="Lab", name="Pearson\nCorrelation") + xlab("") + ylab("")+ geom_tile() + theme_minimal() +  theme(axis.text.x=element_text(angle=45, vjust=1, size=10, hjust=1), axis.text.y = element_text(size=10))


```

The maxium R value is `r round(cormat[maxi,"value"],2)` for variables "`r cormat[maxi,"Var1"]`"" and "`r cormat[maxi,"Var2"]`".

We can plot the different variables against the response variable "dGow"

```{r}
p1<-ggplot(data=data, aes(x=Weight, y=Partition)) + geom_point() 
p2<-ggplot(data=data, aes(x=Heavies, y=Partition)) + geom_point()
p3<-ggplot(data=data, aes(x=Carbons, y=Partition)) + geom_point() 
p4<-ggplot(data=data, aes(x=Polar, y=Partition)) + geom_point() 
p5<-ggplot(data=data, aes(x=Rings, y=Partition)) + geom_point()
p6<-ggplot(data=data, aes(x=HDonors, y=Partition)) + geom_point() 
p7<-ggplot(data=data, aes(x=HAcceptors, y=Partition)) + geom_point() 
p8<-ggplot(data=data, aes(x=GI, y=Partition)) + geom_point() 
library(grid)
library(gridExtra)
grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, ncol=3)
```

And some boxplots

```{r}
par(mfrow=c(1,2))
boxplot(data$Heavies, ylab="Heavies")
boxplot(data$Polar, ylab="Polar")
```


Let's try making a simple OLS using all the variables

```{r}
mm<-lm(Partition~Weight+Heavies+Carbons+Polar+Rings+HDonors+HAcceptors+GI,data=data)
summary(mm)
```

And plot the summary graphs

```{r}
par(mfrow=c(2,2))
plot(mm)
```


Let's do some transforms so we can fix the constant variance 

```{r}
data2<-data
data2$Weight<-log10(data$Weight)
data2[data$Heavies>15,]$Heavies<-15
data2[data$Polar>8,]$Polar<-8
data2$HDonors<-data$HDonors
data2$GI<-log10(data$GI)
p1<-ggplot(data=data2, aes(x=Weight, y=Partition)) + geom_point() 
p2<-ggplot(data=data2, aes(x=Heavies, y=Partition)) + geom_point()
p3<-ggplot(data=data2, aes(x=Carbons, y=Partition)) + geom_point() 
p4<-ggplot(data=data2, aes(x=Polar, y=Partition)) + geom_point() 
p5<-ggplot(data=data2, aes(x=Rings, y=Partition)) + geom_point()
p6<-ggplot(data=data2, aes(x=HDonors, y=Partition)) + geom_point() 
p7<-ggplot(data=data2, aes(x=HAcceptors, y=Partition)) + geom_point() 
p8<-ggplot(data=data2, aes(x=GI, y=Partition)) + geom_point() 
grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, ncol=3)
```

And redo the OLS modelling

```{r}
mm<-lm(Partition~Weight+Heavies+Carbons+Polar+Rings+HDonors+HAcceptors+GI,data=data2)
summary(mm)
par(mfrow=c(2,2))
plot(mm)
```

Now it is time to try data selection - which explanatory variables are most interesting? We will only use backward selection here

```{r}
mm<-lm(Partition~Weight+Heavies+Carbons+Polar+Rings+HDonors+HAcceptors+GI,data=data2)
mm_stepped<-step(mm,trace=T)
summary(mm_stepped)
```

Plot the usualy diagnostic plots

```{r}
par(mfrow=c(2,2))
plot(mm_stepped)
```


Now validate the model on the Minnesota set

```{r}
sota<-read_excel("../minnesota.xlsx")
sota$Weight<-log10(sota$Weight)
sota$GI<-log10(sota$GI)
f<-sota[c("Partition")]
f$pred<-predict(mm_stepped, sota[c("Weight","Carbons","HDonors","HAcceptors","GI")])
```

Some simple statistics of this model is MAD=`r round(mean(abs(f$Partition-f$pred)),2)` kcal/mol and $R$=`r round(cor(f$Partition,f$pred),2)`.

and a correlation graph looks like this 

```{r}
library(ggplot2)
qplot(Partition, pred, data=f, xlab="Experimental [kcal/mol]", ylab="Predicted [kcal/mol]", xlim=c(-5,8), ylim=c(-5,8)) + theme(aspect.ratio=1)
```


