---
title: "An OLS model of water/octanol partition coefficients"
output: html_notebook
---

First we read the data from the Excel sheet and look what data we have

```{r}
library(readxl)
data <- read_excel("../data.xlsx")
colnames(data)
```

The "Nr", "Smiles", "CAS" and "dGow" was taken from the SI of a published paper, and the other columns were filled in with a Python script.

Let's look into the colinearity between the molecular properties

```{r}
library(ggplot2)
library(reshape2)

cormat<-melt(cor(data[,seq(5,dim(data)[2])]))

ii<-which(cormat[,"value"]==1)
maxv<-max(cormat[-ii,"value"])
maxi<-which(cormat["value"]==maxv)[1]

ggplot(data = cormat, aes(x=Var1, y=Var2, fill=value)) + scale_fill_gradient2(low="green", high="red", mid="white", midpoint=0, limit=c(-1,1), space="Lab", name="Pearson\nCorrelation") + xlab("") + ylab("")+ geom_tile() + theme_minimal() +  theme(axis.text.x=element_text(angle=45, vjust=1, size=10, hjust=1), axis.text.y = element_text(size=10))


```

The maxium R value is `r round(cormat[maxi,"value"],2)` for variables "`r cormat[maxi,"Var1"]`"" and "`r cormat[maxi,"Var2"]`".

We can plot the different variables against the response variable "dGow"

```{r}
p1<-ggplot(data=data, aes(x=Weight, y=dGow)) + geom_point() 
p2<-ggplot(data=data, aes(x=Heavies, y=dGow)) + geom_point()
p3<-ggplot(data=data, aes(x=Carbons, y=dGow)) + geom_point() 
p4<-ggplot(data=data, aes(x=Polar, y=dGow)) + geom_point() 
p5<-ggplot(data=data, aes(x=Rings, y=dGow)) + geom_point()
p6<-ggplot(data=data, aes(x=HDonors, y=dGow)) + geom_point() 
p7<-ggplot(data=data, aes(x=HAcceptors, y=dGow)) + geom_point() 
library(grid)
library(gridExtra)
grid.arrange(p1, p2, p3, p4, p5, p6, p7, ncol=3)
```


Let's try making a simple OLS using all the variables

```{r}
mm<-lm(dGow~Weight+Heavies+Carbons+Polar+Rings+HDonors+HAcceptors,data=data)
summary(mm)
```

